Open-source landscape for 3D CSEM modelling

### 2020-06-28 manuscript/Manuscript.tex
first pass at the SimPEG section

@@

The four codes under consideration are, in alphabetical order, \custem
\citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem
\citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg
\citep{CAG.15.Cockett}. All four codes have their user-facing routines written
in Python; all of them make heavy use of \texttt{NumPy}
\citep{CSE.11.VanDerWalt} and \texttt{SciPy} \citep{NM.20.Virtanen}. The four
of them are “modern” open-source projects, meaning that they come with both, an
open-source license and an online-hosted version-control system with tracking
possibilities (raising issues, filing pull requests). All developments comprise
an extensive online documentation with many examples and have continuous
integration to some degree. Newer package-management systems such as Conda,
Docker, or pip significantly supported us to make our codes easily available
and installable without any user-side compilations. Each one can be downloaded
and installed with a single command.

In the following a quick rundown of the codes. It is, however, beyond the scope
of this article to go into every detail of the different modellers, and we
refer to their documentations for more details. An overview comparison of the
different codes is given in Table~\ref{tbl:codecomparison}. All codes have in
common that they solve the\mycom[DW]{are we all using the weak form?} weak
formulation of Maxwell's equation in its differential form (DE) under the
quasistatic or\mycom[DW]{are we all using the diff. approx?}\mycom[OCR]{petgem
consider the weak work of maxwell equations in their diffusive form
(quasi-static approximation)} diffusive approximation, hence neglecting
displacement currents. The machines on which the different codes were run are
listed in Table~\ref{tbl:machines}, together with the responsible operator.

@@

TODO

\clearpage  % TODO: Just for now while drafting, remove later!
\section{Numerical Verifications}

>>>>

\imycom[LH]{Just a thought, but right now we don't state Maxwell's equations anywhere! That seems like something we should do :) That also provides an easy space for talking about Fourier Transform conventions and such too.

% Following the conventions of Ward and Hohmann \cite{Ward1988}, the quasi-static Maxwell's equations are

% \begin{equation}
% \begin{split}
%     \nabla \times \vec{E} &= -i \omega\vec{B}  \\
%     \nabla \times \vec{H} &= \vec{J}
% \end{split}
% \label{eq:maxwell_freq_quasi}
% \end{equation}
}

@@

The four codes under consideration are, in alphabetical order, \custem
\citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem
\citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg
\citep{CAG.15.Cockett, Heagy2017}. All four codes have their user-facing routines written
in Python; all of them make heavy use of \texttt{NumPy}
\citep{CSE.11.VanDerWalt} and \texttt{SciPy} \citep{NM.20.Virtanen}. The four
of them are “modern” open-source projects, meaning that they come with both, an
open-source license and an online-hosted version-control system with tracking
possibilities (raising issues, filing pull requests). All developments comprise
an extensive online documentation with many examples and have continuous
integration to some degree. Newer package-management systems such as Conda,
Docker, or pip significantly supported us to make our codes easily available
and installable without any user-side compilations. Each one can be downloaded
and installed with a single command.

In the following, we provide a quick overview of the codes. It is, however, beyond the scope
of this article to go into every detail of the different modellers, and we
refer to their documentations for more details. An overview comparison of the
different codes is given in Table~\ref{tbl:codecomparison}. All codes have in
common that they solve the\mycom[DW]{are we all using the weak form?}\mycom[LH]{SimPEG uses weak form} weak
formulation of Maxwell's equation in its differential form (DE) under the
quasistatic or\mycom[DW]{are we all using the diff. approx?}\mycom[OCR]{petgem
consider the weak work of maxwell equations in their diffusive form
(quasi-static approximation)}\mycom[LH]{SimPEG uses quasistatic approx.} diffusive approximation, hence neglecting
displacement currents. The machines on which the different codes were run are
listed in Table~\ref{tbl:machines}\mycom[LH]{SimPEG also interfaces to the scipy solvers (but in practice those are use much less than Pardiso / mumps)}, together with the responsible operator.

@@

\simpeg is a modular toolbox for simulations and gradient based inversions in
geophysics. Current functionality includes modelling and inversion
capabilities for gravity, magnetics, direct current resistivity, induced
polarization, electromagnetics (time, frequency, as well as natural sources,
e.g. magnetotellurics), and fluid flow (Richards equation). It is a community
driven project that aims to support researchers by providing a flexible,
extensible framework and to facilitate integration of data sets from a variety
of geophysical methods, including joint inversions \citep{Astic2020}, by
providing a common interface to each method.

Meshes and finite volume differential operators are implemented in the
\texttt{discretize} package. It currently includes tensor, OcTree,
cylindrical, and logically rectangular meshes. Each mesh type inherits from a
common structure and uses the same naming conventions for methods and
properties. This allows us to decouple the implementation of a discretized
partial differential equation from the details of the mesh geometry and
therefore we can write a single implementation of discretized partial
differential equation (or set of partial differential equations) in \simpeg
that will support all mesh types implemented in \texttt{discretize}.

The electromagnetics implementations use a staggered grid finite volume
approach where physical properties are discretized at cell centers, fields on
cell edges and fluxes on cell faces \citep{IEEE.66.Yee} (for an overview of
discretizing Maxwell's equations, we recommend \cite{Haber2014}). There are
two different discretization strategies implemented for Maxwell's equations: (
a) the EB-formulation, which discretizes the electric field ($\vec{e}$) on
cell edges and the magnetic flux density ($\vec{h}$) on cell faces, and (b)
the HJ-formulation, which discretizes the magnetic field ($\vec{h}$) on cell
edges and the current density ($\vec{j}$) on cell faces. In both cases the
physical properties: electrical conductivity / resistivity and magnetic
permeability are discretized at cell centers. Having multiple implementations
allows for testing that compares results from each approach, as well as the
representation of both electrical and magnetic sources on cylindrically
symmetric meshes. \simpeg supports variable magnetic permeability and
full-tensor anisotropy for the physical properties.

% Support for domain decomposition approaches (see for example \cite{Fournier2020})

\simpeg interfaces to various solvers including \texttt{PARDISO} \citep{FGCS.04.Schenk}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, as well as those implemented in the \texttt{scipy.sparse.linalg} module \citep{NM.20.Virtanen}. For problems with many sources (and reciprocity is not a natural solution, e.g. in 3D airborne EM), \simpeg contains machinery for domain-decomposition approaches; see for example \cite{Fournier2020}.

\clearpage  % TODO: Just for now while drafting, remove late{}r!{}
\section{Numerical Verifications}
